---
title: "Biometry HW06"
author: "Joel Yeager"
date: "2024-10-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
```

## Section 7.1

1.  Compute the correlation coefficient for the following data set.

    x \| 1 2 3 4 5 6 7\
    y \| 2 1 4 3 7 5 6

```{r 7.1.1}
x <- c(1:7)
y <- c(2,1,4,3,7,5,6)

corr <- cor(x, y)
corr
```

3.  For each of the following scatterplots, state whether the correlation coefficient is an appropriate summary, and explain briefly.

    a.  Correlation coefficient is seemingly appropriate because data appears linear,\
        and shows positive correlation.
    b.  Correlation coefficient is not appropriate as the data do not appear\
        to be linearly associated.
    c.  Correlation coeficient not appropriate as there are several outliers\
        present in the data.

4.  An investigator collected data on heights and weights of college students.\
    The correlation between height and weight for men was about 0.6, and for\
    women it was about the same. If men and women are taken together, will the\
    correlation between height and weight be more than 0.6, less than 0.6,\
    or about equal to 0.6? It might be helpful to make a rough scatterplot.

    Correlation coefficient would be less than 0.6 as the additional\
    between group variation remains unexplained by simple x-y relationships.

## Section 7.2

1.  Each month for several months, the average temperature in C (x) and the number of pounds of steam (y) consumed by a certain chemical plant were measured. The least-squares line computed from the resulting data is y = 245.82 + 1.13x.

    a.  Predict the number of pounds of steam consumed in a month where the\
        average temperature is 65C.

```{r 7.2.1a }
(p7.2.1a <- 245.82 + (1.13 * 65))
```

```         
b.  If two months differ in their average temperatures by 5C, by how much\
    do you predict the number of pounds of steam consumed to differ?
```

```{r 7.2.1b }
(p7.2.1b <- 1.13 * 5)
```

2.  A least-squares line is fit to a set of points. If the total sum of squares\
    is SUM(y~i~ - ybar)\^2 = 9615, and the error sum of squares is\
    SU\<(y~i~ - yhat~i~)\^2 = 1450, compute the coefficient of determination r\^2.

```{r 7.2.3}
(r2 <- 1450 / 9615)
```

5.  In Galton’s height data (Figure 7.1, in Section 7.1), the least-squares line\
    for predicting forearm length (y) from height (x) is y = -0.2967 + 0.2738x.

    a.  Predict the forearm length of a man whose height is 70 in.

```{r 7.2.5a}
(p7.2.5a <- -0.2967 + (0.2738 * 70))
```

```         
b.  How tall must a man be so that we would predict his forearm length to be 19 in.?
```

```{r 7.2.5b}
(p7.2.5b <- (19 + 0.2967) / 0.2738)
```

```         
c.  All the men in a certain group have heights greater than the height\
    computed in part (b). Can you conclude that all their forearms will be at\
    least 19 in. long? Explain.\
```

No. Observations below the least-squares line would have forearms \< 19 in.

6.  Moisture content in percent by volume (x) and conductivity in mS/m (y) were\
    measured for 50 soil specimens. The means and standard deviations were x = 8.1,\
    sx = 1.2, y = 30.4, sy = 1.9. The correlation between conductivity and moisture\
    was computed to be r = 0.85. Find the equation of the least-squares line\
    for predicting soil conductivity from moisture content.

```{r 7.2.7}
(b.1 <- 1.9 * 0.85 / 1.2)

(b.0 <- 30.4 - b.1 * 8.1)

# plugged into "y = mx + b"; y = 1.345833x + 19.49875
```

## Section 7.4

1.  The following output (from MINITAB) is for the least-squares fit of the\
    model ln y = Beta0 + Beta1 ln x + E, where y represents the monthly production\
    of a gas well and x represents the volume of fracture fluid pumped in.\
    (A scatterplot of these data is presented in Figure 7.22.)

    a.  What is the equation of the least-squares line for predicting ln y from ln x?\
        **ln(y) = -0.4442 + 0.79833 \* ln(x)**

    b.  Predict the production of a well into which 2500 gal/ft of fluid have been pumped.

```{r 7.4.1b}
(p7.4.1b <- exp(-0.4442 +0.79833 * log(2500)))
```

```         
c.  Predict the production of a well into which 1600 gal/ft of fluid have been pumped.
```

```{r 7.4.1c}
(p7.4.1c <- exp(-0.4442 + 0.79833 * log(1600)))
```

```         
d.  Find a 95% prediction interval for the production of a well into which\
    1600 gal/ft of fluid have been pumped.(Note: ln 1600 = 7.3778.)
```

```{r 7.4.1d}
(LCI <- exp(3.9738))
(UCI <- exp(6.9176))
```
