---
title: "BiometryHW02"
output: pdf_document
date: "2024-09-04"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rmarkdown)
library(magrittr)
library(gtools)
```

## Biometry HW02 - Yeager

### Section 2.1

1.  The probability that a bearing fails during the first month of use
    is 0.12. What is the probability that it does not fail during the
    first month?

```{r s2.1 q1}
p.fail <- 0.12
p.succ <- 1 - p.fail
p.succ
```

3.  A section of an exam contains four True-False questions. A completed
    exam paper is selected at random, and the four answers are
    recorded.\

    A. List all 16 outcomes in the sample space.

```{r s2.1 q3.a}}
q.num <- c('Q1','Q2','Q3','Q4')
outcomes <- expand.grid(lapply(q.num, function(x) c("T","F")))
colnames(outcomes) <- q.num
outcomes
```

```         
B. Assuming the outcomes to be equally likely, find the probability that
all the answers are the same.
```

```{r s2.1 q3.b}
pr.3 <- 1/16 # probability for any one outcome

pr.all.T <- pr.3 # probability all are TRUE
pr.all.F <- pr.3 # probability all are FALSE

# Answers can only be TRUE or FALSE, thus only two of 16 outcomes can have same answers
(pr.same <- pr.all.T + pr.all.F) 
```

```         
C. Assuming the outcomes to be equally likely, find the probability that
exactly one of the four answers is “True.”
```

```{r s2.1 q3.c}
q1.T <- 1/16 # each of 16 outcomes equally likely
q2.T <- 1/16
q3.T <- 1/16
q4.T <- 1/16
pr.one.T <- q1.T + q2.T + q3.T + q4.T # total of 4 outcomes where only 1 is "TRUE"
pr.one.T
```

```         
D. Assuming the outcomes to be equally likely, find the probability that
at most one of the four answers is “True.”
```

```{r s2.1 q3.d}
# There are 5 equally probable outcomes with "TRUE" as the answer for 1 or fewer questions
pr.3 <- 1/16 # probability for any one outcome
pr.max1.T <- pr.3 + pr.3 + pr.3 + pr.3 + pr.3
pr.max1.T
```

5.  Four candidates are to be interviewed for a job. Two of them,
    numbered 1 and 2, are qualified, and the other two, numbered 3 and
    4, are not. The candidates are interviewed at random, and the first
    qualified candidate interviewed will be hired. The outcomes are the
    sequences of candidates that are interviewed. So one outcome is 2,
    and another is 431.

```         
A. List all the possible outcomes.
```

```{r s2.1 q5.a}
library(gtools)
candidates <- c(1, 2, 3, 4)
all.perms <- permutations(n = length(candidates), r = length(candidates), v = candidates)
colnames(all.perms) <- c('first','second','third','fourth')
all.perms %<>% as.tibble()
all.perms$sequence <- paste0("Seq.", 1:24)
all.perms
```

```         
B. Let A be the event that only one candidate is interviewed. List the outcomes in "A".
```

```{r s2.1 q5.b}
A <- all.perms %>% filter(first %in% c(1,2))
A

# Interviews stop after candidate 1 or 2, and only one interview is conducted, so although other sequences exist there are only two outcomes; either candidate 1 or candidate 2 are interviewed first.
```

```         
C. Let B be the event that three candidates are interviewed. List the outcomes in B.
```

```{r s2.1 q5.c}
B <- all.perms %>% filter(!first %in% c(1,2)) %>% filter(!second %in% c(1,2))
B

# for three candidates to be interviewed, candidates 1 & 2 must be interviewed third or fourth
```

```         
D. Let C be the event that candidate 3 is interviewed. List the outcomes in C.
```

```{r s2.1 q5.d}
C <- all.perms %>% filter(first %in% c(3,4)) %>% 
  filter(!(first == 4 & second != 3))
C

# Outcomes where Candidate 3 are interviewed do not have Candidates 1 or 2 preceed them.
```

```         
E. Let D be the event that candidate 2 is not interviewed. List the outcomes in D.
```

```{r s2.1 q5.e}
D1 <- all.perms %>% filter(first == 1)
D2 <- all.perms %>% filter(second == 1 & first != 2)
D3 <- all.perms %>% filter(third == 1 & first != 2 & second != 2)
D <- rbind(D1,D2,D3)
D
# Outcomes where candidate 1 is interviewed before candidate 2 result in the event that candidate 2 is not interviewed.
```

```         
F. Let E be the event that candidate 4 is interviewed. Are A and E mutually exclusive? Howabout B and E, C and E, D and E?
```

```{r s2.1 q5.f}
E <- all.perms %>% filter(first == 4 | second == 4) %>% filter(!first %in% c(1,2))
E
```

7.  In a survey of households with television sets, the proportion of
    television sets in various types of rooms was

```{r s2.1 q7}
room <- c('bed','liv','den','bas','kit','bat')
tv <- c(0.37,0.26,0.22,0.12,0.02,0.01)
d <- tibble(room, tv)
d
```

```         
A. What is the probability that a TV set is located in a living room or den?
```

```{r s2.1 q7.a}
pr.liv <- 0.26
pr.den <- 0.22
pr.either <- pr.liv + pr.den
pr.either

# The event is not mutually exclusive, thus the probability of a TV being in either the living room or the den is equal to the sum of the probabilities
```

```         
B. What is the probability that a TV set is not located in a bedroom?
```

```{r s2.1 q7.b}
pr.bed <- 0.37
pr.not.bed <- 1 - pr.bed
pr.not.bed
```

9.  Among the cast aluminum parts manufactured on a certain day, 80%
    were flawless, 15% had only minor flaws, and 5% had major flaws.
    Find the probability that a randomly chosen part:

```{r s2.1 q9}
flawless <- 0.80
minor <- 0.15
major <- 0.05
```

```         
A. has a flaw (major or minor).
```

```{r s2.1 q9.a}
pr.flawed <- major + minor
pr.flawed
```

```         
B. has no major flaw.
```

```{r s2.1 q9.b}
pr.no.major <- 1 - major
pr.no.major
```

11. A quality-control engineer samples 100 items manufactured by a
    certain process, and finds that 15 of them are defective. True or
    false:

```         
A. The probability that an item produced by this process is
   defective is 0.15.
```

***FALSE***

```         
B. The probability that an item produced by this process is 
   defective is likely to be close to 0.15, 
   but not exactly equal to 0.15.
```

***TRUE***

13. Let S be the event that a randomly selected college student has
    taken a statistics course, and let C be the event that the same
    student has taken a chemistry course. Suppose P(S) = 0.4, P(C) =
    0.3, and P(S ∩ C) = 0.2.

```{r s2.1 q13}
P.S <- 0.4
P.C <- 0.3
P.intersect <- 0.2
```

```         
A. Find the probability that a student has taken statistics, chemistry,
or both.
```

```{r s2.1 q13.a}
P.SUC <- P.S + P.C - P.intersect
P.SUC
```

```         
B. Find the probability that a student has taken neither statistics nor
chemistry.
```

```{r s2.1 q13.b}
P.neither <- 1 - P.SUC
P.neither
```

```         
C. Find the probability that a student has taken statistics but not
chemistry.
```

```{r s2.1 q13.c}
P.SnC <- P.S + P.C - P.SUC
P.SnC
```

15. All the fourth-graders in a certain elementary school took a
    standardized test. A total of 85% of the students were found to be
    proficient in reading,78% were found to be proficient in
    mathematics, and 65% were found to be proficient in both reading and
    mathematics. A student is chosen at random.

```{r s2.1 q15}
reading <- 0.85
math <- 0.78
both <- 0.65
```

```         
A. What is the probability that the student is proficient
   in mathematics but not in reading?
```

```{r s2.1 q15.a}
P.RCnM <- math - both
P.RCnM
```

```         
B. What is the probability that the student is proficient
   in reading but not in mathematics?
```

```{r s2.1 q15.b}
P.RnMC <- reading - both
P.RnMC
```

```         
C. What is the probability that the student is proficient
   in neither reading nor mathematics?
```

```{r s2.1 q15.c}
P.RUM <- reading + math - both
P.RCUMC <- 1 - P.RUM
P.RCUMC
```

17. A system contains two components, A and B. The system will function
    only if both components function. The probability that A functions
    is 0.98, the probability that B functions is 0.95, and the
    probability that either A or B functions is 0.99. What is the
    probability that the system functions?

```{r s2.1 q17}
P.A <- 0.98
P.B <- 0.95
P.AUB <- 0.99
P.AnB <- P.A + P.B - P.AUB
P.AnB
```

19. True or false: If A and B are mutually exclusive,

```         
A. P(A ∪ B) = 0
```

*FALSE*

```         
B. P(A ∩ B) = 0
```

*TRUE*

```         
C. P(A ∪ B) = P(A ∩ B)
```

*FALSE*

```         
D. P(A ∪ B) = P(A) + P(B)
```

*TRUE*

### Section 2.2

1.  DNA molecules consist of chemically linked sequences of the bases
    adenine, guanine, cytosine, and thymine, denoted A, G, C, and T. A
    sequence of three bases is called a codon. A base may appear more
    than once in a codon.

```         
A. How many different codons are there?
```
```{r s2.2 q1.a1}
# to simply get number of codons:
num.codons <- 4 * 4 * 4
num.codons
```

```{r s2.2 q1.a2}
# to see all codons 
bases <- c('A','G','C','T')
codons <- permutations(n = length(bases), r = 3, 
                       v = bases, repeats.allowed = TRUE)
codons <- codons %>% as_tibble()
head(codons)
```
```         
B. The bases A and G are purines, while C and T are
   pyrimidines. How many codons are there whose
   first and third bases are purines and whose second
   base is a pyrimidine?
```
```{r s2.2 q1.b}
q1.b <- 2 * 2 * 2
q1.b
```
```{r s2.2 q1.b2}
# to list out codon sequences:
q1.b2 <- codons %>% filter(V1 %in% c('A','G') & V3 %in% c('A','G')) %>% 
  filter(V2 %in% c('C','T'))
q1.b2
```
```
C. How many codons consist of three different bases?
```
```{r s2.2 q1.c}
q1.c <- 
```