---
title: "HW04_Yeager"
output: pdf_document
date: "2024-10-01"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(magrittr)
```

## Section 5.1

1.  Find the value of z(alpha/2) to use in expression (5.1) to construct
    a confidence interval with level:

```{r 5.1.1}
## a) 95%
# using z-score table, 1.96

## b) 98%
# using z-score table, 2.33

## c) 99%
# using z-score table, 2.58

## d) 80%
# using z-score table, 1.28 (or 1.29)
```

3.  As the confidence level goes up, the reliability goes \_\_\_\_\_\_\_
    and the precision goes \_\_\_\_\_\_\_.

    -   First blank: up; second blank: down.

4.  In a sample of 100 steel wires the average breaking strength is 50
    kN, with a standard deviation of 2 kN.

```{r 5.1.5}
## a) Find a 95% confidence interval for the mean 
##    breaking strength of this type of wire.
n <- 100
x.bar <- 50
sd <- 2

(a1 <- x.bar + (1.96 * (2 / sqrt(n))))
(a2 <- x.bar - (1.96 * (2 / sqrt(n))))
# 95% CI: 49.608 to 50.392

## b) Find a 99% confidence interval for the mean
##    breaking strength of this type of wire.
(b1 <- x.bar + (2.58 * (sd / sqrt(n))))
(b2 <- x.bar - (2.58 * (sd / sqrt(n))))
# 99% CI: 49.484 to 50.516

## c) An engineer claims that the mean breaking
##    strength is between 49.7 kN and 50.3 kN. With
##    what level of confidence can this statement be
##    made?
(z <- (50.3 - 50) / (2 / sqrt(n)))
(alp.2 <- 1 - 0.9332)
(conf <- (1 - (2 * alp.2)) * 100)

## d) How many wires must be sampled so that a 95%
#     confidence interval specifies the mean breaking
#     strength to within +/- 0.3 kN?
z.025 <- 1.96
(two.sqrt.n <- (0.3 / z.025))
(n <- (2 / two.sqrt.n)^2)
# n = 171

## e) How many wires must be sampled so that a 99%
#     confidence interval specifies the mean breaking
#     strength to within +/- 0.3 kN?
z.005 <- 2.58
(two.sqrt.n <- (0.3 / z.005))
(n <- (2 / two.sqrt.n)^2)
# n = 296
```

7.  The capacities (in ampere-hours) were measured for a sample of 120
    batteries. The average was 178 and the standard deviation was 14.

```{r 5.1.7}
## a) Find a 95% confidence interval for the mean capacity
#     of batteries produced by this method.
n <- 120
x.bar <- 178
sd <- 14

(a1 <- x.bar + (1.96 * (sd / sqrt(n))))
(a2 <- x.bar - (1.96 * (sd / sqrt(n))))
# 95% CI: 175.4951 to 180.5049

## b) Find a 99% confidence interval for the mean capacity
#     of batteries produced by this method.
(b1 <- x.bar + (2.58 * (sd / sqrt(n))))
(b2 <- x.bar - (2.58 * (sd / sqrt(n))))
# 99% CI: 174.7027 to 181.2973

## c) An engineer claims that the mean capacity is
#     between 176 and 180 ampere-hours. With what
#     level of confidence can this statement be made?
(z <- (180 - x.bar) / (sd / sqrt(n))) # 1.56
(alp.2 <- 1 - 0.9406)
(conf <- (1 - (2 * alp.2)) * 100)

## d) Approximately how many batteries must be sampled
#     so that a 95%confidence interval will specify
#     the mean to within +/- 2 ampere-hours?
z.025 <- 1.96
(sd.sqrt.n <- (2 / z.025))
(n <- (sd / sd.sqrt.n)^2)
# 189 batteries must be sampled

## e) Approximately how many batteries must be sampled
#     so that a 99%confidence interval will specify
#     the mean to within  +/- 2 ampere-hours?
z.005 <- 2.58
(sd.sqrt.n <- (2 / z.005))
(n <- (sd / sd.sqrt.n)^2)
# 327 batteries
```

## Section 5.3

1.  Find the value of t(n-1,alpha/2) needed to construct a two-sided
    confidence interval of the given level with the given sample size:

```{r 5.3.1}
## a) Level 90%, sample size 12.
A.n <- 12
(A.df <- A.n - 1)
(A.alpha.2 <- (1 - .9) / 2)
# df = 11, alpha/2 = 0.05; t = 1.796

## b) Level 95%, sample size 7.
B.n <- 7
(B.df <- B.n - 1)
(B.alpha.2 <- (1 - .95) / 2)
# df = 6, alpha/2 = 0.025; t = 2.447

## c) Level 99%, sample size 2.
C.n <- 2
(C.df <- C.n - 1)
(C.alpha.2 <- (1 - .99) / 2)
# df = 1, alpha/2 = 0.005; t = 63.657

## d) Level 95%, sample size 29.
D.n <- 29
(D.df <- D.n - 1)
(D.alpha.2 <- (1 - .95) / 2)
# df = 28, alpha/2 = 0.025; t = 2.048
```

3.  Find the level of a two-sided confidence interval that is based on
    the given value of t(n-1,alpha/2) and the given sample size.

```{r 5.3.3}
## a. t = 2.776, sample size 5.
A.n <- 5
(A.df <- A.n - 1)
A.alpha <- 0.025 # from table
(A.CI <- (1 - (2 * A.alpha)) * 100)
# 95% CI

## b. t = 2.718, sample size 12.
B.n <- 12
(B.df <- B.n - 1) # 11
B.alpha <- 0.01 # from table
(B.CI <- (1 - (2 * B.alpha)) * 100)
# 98% CI

## c. t = 5.841, sample size 4.
C.n <- 4
(C.df <- C.n - 1) # 3
C.alpha <- 0.005 # from table
(C.CI <- (1 - (2 * C.alpha)) * 100)
# 99% CI

## d. t = 1.325, sample size 21.
D.n <- 21
(D.df <- D.n - 1) # 20
D.alpha <- 0.1 # from table
(D.CI <- (1 - (2 * D.alpha)) * 100)
# 80% CI

## e. t = 1.746, sample size 17.
E.n <- 17
(E.df <- E.n - 1) # 16
E.alpha <- 0.05 # from table
(E.CI <- (1 - (2 * E.alpha)) * 100)
# 90% CI
```

5.  The article “Wind-Uplift Capacity of Residential Wood Roof-Sheathing
    Panels Retrofitted with Insulating Foam Adhesive” (P. Datin, D.
    Prevatt, and W. Pang, Journal of Architectural Engineering,
    2011:144–154) presents a study of the failure pressures of roof
    panels. Following are the failure pressures, in kPa, for five panels
    constructed with 6d smooth shank nails. These data are consistent
    with means and standard deviations presented in the article.

    3.32 2.53 3.45 2.38 3.01

    Find a 95% confidence interval for the mean failure pressure for
    this type of roof panel.

```{r 5.3.5}
d <- c(3.32, 2.53, 3.45, 2.38, 3.01)
(x.bar <- mean(d))
(s <- sd(d))
n <- 5
(df <- n - 1) # 4
t <- 2.776 # from table
(u.CI <- x.bar + (t * (s / sqrt(n))))
(l.CI <- x.bar - (t * (s / sqrt(n))))
# (2.352047, 3.523953)
```

7.  The article “An Automatic Visual System for Marble Tile
    Classification” (L. Carrino, W. Polini, and S. Turchetta, Journal of
    Engineering Manufacture, 2002:1095–1108) describes a measure for the
    shade of marble tile in which the amount of light reflected by the
    tile is measured on a scale of 0–255. A perfectly black tile would
    reflect no light and measure 0, and a perfectly white tile would
    measure 255. A sample of nine Mezza Perla tiles were measured, with
    the following results: 204.999 206.149 202.102 207.048 203.496
    206.343 203.496 206.676 205.831 Is it appropriate to use the
    Student’s t statistic to construct a 95% confidence interval for the
    mean shade of Mezza Perla tile? If so, construct the confidence
    interval. If not, explain why not.

```{r 5.3.7}
tiles <- c(204.999, 206.149, 202.102, 207.048, 203.496,
           206.343, 203.496, 206.676, 205.831)
shapiro.test(tiles)
# Yes, the test is appropriate as the data appear to be normally distributed
# as W = 0.90314 is close to W = 1, and a p-value > 0.05 (p-value 0.2707)
# indicates these data are not significantly different 
# from a normal distribution.

(x.bar <- mean(tiles))
(s <- sd(tiles))
(n <- length(tiles))
(df <- n - 1)
t8.025 <- 2.306 # from table
(u.CI <- x.bar + (t8.025 * (s/sqrt(n))))
(l.CI <- x.bar - (t8.025 * (s/sqrt(n))))
# (203.8066, 206.4468) or 205.1267 +/- 1.3201
```
