---
title: "Biometry HW03"
author: "Joel Yeager"
date: "2024-09-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 4.1

1.  After scoring a touchdown, a football team may elect to attempt a two-point conversion, by running or passing the ball into the end zone. If successful, the team scores two points. For a certain football team, the probability that this play is successful is 0.40.

    -   A:

    ```{r s4.1.1a}
    # Let X = 1 if successful, X = 0 if not. Find the mean and variance of X.
    p <- 0.4
    (mu.X <- p) # mean of X is the same as Pr(Success)
    (sigma.X <- p * (1 - p))
    ```

    -   B: no because the possible values for Y are not 0 and 1

    -   C:

    ```{r s4.1.1c}
    (mu.Y <- 2 * p) # because is 2 bernoulli's
    (sigma.Y <- (4 * p) * (1 - p)) # or 2^2 * sigma.X^2
    ```

3.  When a certain glaze is applied to a ceramic surface, the probability is 5% that there will be discoloration, 20% that there will be a crack, and 23% that there will be either discoloration or a crack, or both. Let X = 1 if there is discoloration, and let X = 0 otherwise. Let Y = 1 if there is a crack, and let Y = 0 otherwise.

```{r s4.1.3}
p.disc <- 0.05
p.crack <- 0.2
p.either <- 0.23

#(a)
(px <- p.disc) # px is same as probability for discoloration
#(b)
(py <- p.crack) # py here is same as crack probability
#(c)
(pz <- p.either) # pz is same as probability either defect being present
#(d)
# Yes, the defects are not mutually exclusive.
#(e)
# No because pz can be zero while px + py cannot.
#(f)
# No, because Z must equal 1 where X + Y = 2
```
## Section 4.2

Q1. Let X ~ Bin(7, 0.3) Find:
```{r s4.2.1}
# define necessary variables first
n <- 7
p <- 0.3

# a) P(X = 1)
X.a <- 1
(a <- (factorial(n))/(factorial(X.a)*factorial(n - X.a))*p^X.a*(1 - p)^(n-X.a))

# b) P(X = 2)
X.b <- 2
(b <- (factorial(n))/(factorial(X.b)*factorial(n - X.b))*p^X.b*(1 - p)^(n-X.b))

# c) P(X < 1)
X.c <- 0
(c <- (factorial(n))/(factorial(X.c)*factorial(n - X.c))*p^X.c*(1 - p)^(n-X.c))

# d) P(X > 4)
X.d1 <- 5
X.d2 <- 6
X.d3 <- 7

(d <- (factorial(n))/(factorial(X.d1) * factorial(n-X.d1))*p^X.d1*(1 - p)^(n - X.d1) + (factorial(n))/(factorial(X.d2) * factorial(n - X.d2))*p^X.d2*(1 - p)^(n - X.d2) + (factorial(n))/(factorial(X.d3)*factorial(n - X.d3))*p^X.d3*(1 - p)^(n - X.d3))

# e) mu X
(mu <- n * p)

# f) sigma^2
(sig.sq <- n * p * (1 - p))
```

Q5. Of all the registered automobiles in a certain state, 10% violate the state emissions standard. Twelve automobiles are selected at random to undergo an emissions test.

```{r s4.2.5}
# a) Find the probability that exactly three of them violate the standard.
n <- 12
p <- 0.1

X.a <- 3 # the three that "violate the standard"
(a <- (factorial(n))/(factorial(X.a) * factorial(n - X.a)) * p^X.a * (1 - p)^(n - X.a))

# b) Find the probability that fewer than three of them violate the standard.
xb.1 <- 0
xb.2 <- 1
xb.3 <- 2

b <- (factorial(n))/(factorial(xb.1) * factorial(n - xb.1)) * p^xb.1 * (1 - p)
```