---
title: "Yeager_HW07"
author: "Joel Yeager"
date: "2024-10-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Section 8.1

1.  In an experiment to determine the factors affecting tensile strength in steel plates, the tensile strength (in kg/mm2), the manganese content (in parts per thousand), and the thickness (in mm) were measured for a sample of 20 plates. The following MINITAB output presents the results of fitting the model Tensile strength = B0 + B1 Manganese + B2 Thickness.

    a.  Predict the strength for a specimen that is 10 mm thick and contains\
        8.2 ppt manganese.

    ```{r 8.1.1a}
    (p8.1.1a <- 26.641 + 3.3201 * 8.2 - 0.4249 * 10)
    ```

    b.  If two specimens have the same thickness, and one contains 10 ppt more\
        manganese, by how much would you predict their strengths to differ?

    ```{r 8.1.1b}
    (p8.1.1b <- 3.3201 * 10)
    ```

    c.  If two specimens have the same proportion of manganese, and one is\
        5 mm thicker than the other, by how much would you predict their\
        strengths to differ?

    ```{r 8.1.1c}
    (p8.1.1c <- 0.4249 * 5)
    ```

2.  The data used to fit the model in Exercise 1 are presented in the\
    following table, along with the residuals and the fitted values. Plot the\
    residuals versus the fitted values. Does the plot indicate that the linear\
    model is reasonable? Explain.

```{r 8.1.3}
res<- c(-0.111, -0.709, 0.091, 0.291, -0.517, 0.790, -0.517, -0.010, 0.89,
        0.214, -0.464, -0.864, 0.121, -0.164, -0.181, -0.206, 1.619, -1.764,
        1.446, 0.04)
fit<- c(47.811, 51.609, 51.609, 51.609, 50.517, 49.710, 50.517, 49.710, 49.710,
        47.486, 47.564, 45.864, 47.479, 45.864, 47.181, 45.906, 47.181, 47.564,
        47.054, 48.560)
plot(fit, res)
abline(0,0)
```

No apparent pattern of residuals, thus a non-linear relationship\
is not likely and a linear model would be more appropriate.

5.  a fitted model for predicting the extraction of manganese in % (y) from\
    particle size in mm (x~1~), the amount of sulfur dioxide in multiples of the\
    stoichiometric quantity needed for the dissolution of manganese (x~2~),\
    and the duration of leaching in minutes (x~3~) is given as:

y = 56.145 − 9.046x~1~ − 33.421x~2~ + 0.243x~3~ − 0.5963x~1~x~2~ − 0.0394x~1~x~3~ + 0.6022x~2~x~3~ + 0.6901x~1~^2^ + 11.7244x~2~^2^ − 0.0097x~3~^2^

  There were a total of n = 27 observations, with SSE = 209.55 and SST = 6777.5.

    a. Predict the extraction percent when the particle size is 3 mm,\
    the amount of sulfur dioxide is 1.5, and the duration of leaching\
    is 20 minutes.
```{r 8.1.5a}
(q8.1.5a <- 56.145 - 9.046 * 3 - 33.421 * 1.5 + 0.243 * 20 - 0.5963 * 3 * 1.5 - 0.0394 * 3 * 20 + 0.6022 * 1.5 * 20 + 0.6901 * (3^2) + 11.7244 * (1.5^2) - 0.0097 * (20^2))
```

    b. Is it possible to predict the change in extraction percent when the\
    duration of leaching increases by one minute? If so, find the predicted\
    change. If not, explain why not. 
  No. There are additional factors that influence the relationship.
  
    c. Compute the coefficient of determination *R~2~*.
```{r 8.1.5c}
(q8.1.5c <- (6777.5 - 209.55) / 6777.5)
```

    d. Compute the F statistic for testing the null hypothesis that all the\
    coefficients are equal to 0. Can this hypothesis be rejected?
```{r 8.1.5d}
(fStat <- ((6777.5 - 209.55) / 9) / (209.55 / 17))

# According to table for this F-statistic, p < 0.001 and thus we should reject
# the null hypothesis
```

7. In a study of the lung function of children, the volume of air exhaled under force in one second is called FEV1.Measurements were made on a group of children each year for two years. A linear model was fit to predict this year’s FEV1 as a function of last year’s FEV1 (in liters), the child’s gender (0 = Male, 1 = Female), the child’s height (in m), and the ambient atmospheric pressure (in mm). The following MINITAB output presents the results of fitting the model

    a. Predict the FEV1 for a boy who is 1.4 m tall, if the measurement was\
       taken at a pressure of 730 mm and last year’s measurement was 2.113 L.
```{r 8.1.7a}
(q8.1.7a <- -0.21947 + 0.779 * 2.113 - 0.10827 * 0 + 1.3536 * 1.4 - 0.0013431 * 730)
```

    b. If two girls differ in height by 5 cm, by how much would you expect\
       their FEV1 measurements to differ, other things being equal?
```{r 8.1.7b}
(q8.1.7b <- 1.3536 * 0.05)
```

    c. The constant term B~0~ is estimated to be negative. But FEV1 must always\
    be positive. Is something wrong? Explain.
    
  This is fine, B~0~ is the y-intercept and because height is on the y axis\
  and a person cannot have a negative height the value is theoretical.
  
9. (too much text to include)

    a. Predict the pH for a pulp with a viscosity of 1500 kg/m * s, a density\
       of 1.04 g/cm^3^, and a BRIX of 17.5%.
```{r 8.1.9a}
(pH <- -1.7914 + 0.00026626 * 1500 + 9.8184 * 1.04 - 0.29982 * 17.5)
```

    b. If two pulps differ in density by 0.01 g/cm3, by how much would you\
       expect them to differ in pH, other things being equal?
```{r 8.1.9b}
(q8.1.9b <- 9.8184 * 0.01)
```

    c. The constant term β0 is estimated to be negative. But pulp pH must\
       always be positive. Is something wrong? Explain.
  This is also fine. It is the theoretical pH value for a pulp with density = 0


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
