---
title: "BiometryHW02"
output: pdf_document
date: "2024-09-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rmarkdown)
library(magrittr)
library(gtools)
```

## Biometry HW02 - Yeager

> **Section 2.1**

1.  The probability that a bearing fails during the first month of use is 0.12. What is the probability that it does not fail during the first month?

```{r s2.1 q1}
p.fail <- 0.12
p.succ <- 1 - p.fail
p.succ
```

3.  A section of an exam contains four True-False questions. A completed exam paper is selected at random, and the four answers are recorded.\

A. List all 16 outcomes in the sample space.

```{r s2.1 q2.a}}
q.num <- c('Q1','Q2','Q3','Q4')
outcomes <- expand.grid(lapply(q.num, function(x) c("T","F")))
colnames(outcomes) <- q.num
outcomes
```

B. Assuming the outcomes to be equally likely, find the probability that all the answers are the same.

```{r s2.1 q2.b}
pr.3 <- 1/16 # probability for any one outcome

pr.all.T <- pr.3 # probability all are TRUE
pr.all.F <- pr.3 # probability all are FALSE

# Answers can only be TRUE or FALSE, thus only two of 16 outcomes can have same answers
(pr.same <- pr.all.T + pr.all.F) 
```

C. Assuming the outcomes to be equally likely, find the probability that exactly one of the four answers is “True.”

```{r s2.1 q2.c}
q1.T <- 1/16 # each of 16 outcomes equally likely
q2.T <- 1/16
q3.T <- 1/16
q4.T <- 1/16
pr.one.T <- q1.T + q2.T + q3.T + q4.T # total of 4 outcomes where only 1 is "TRUE"
pr.one.T
```

D. Assuming the outcomes to be equally likely, find the probability that at most one of the four answers is “True.”

```{r s2.1 q2.d}
# There are 5 equally probable outcomes with "TRUE" as the answer for 1 or fewer questions
pr.3 <- 1/16 # probability for any one outcome
pr.max1.T <- pr.3 + pr.3 + pr.3 + pr.3 + pr.3
pr.max1.T
```

5.  Four candidates are to be interviewed for a job. Two of them, numbered 1 and 2, are qualified, and the other two, numbered 3 and 4, are not. The candidates are interviewed at random, and the first qualified candidate interviewed will be hired. The outcomes are the sequences of candidates that are interviewed. So one outcome is 2, and another is 431.

A. List all the possible outcomes.

```{r s2.1 q5.a}
library(gtools)
candidates <- c(1, 2, 3, 4)
all.perms <- permutations(n = length(candidates), r = length(candidates), v = candidates)
colnames(all.perms) <- c('first','second','third','fourth')
all.perms %<>% as.tibble()
all.perms$sequence <- paste0("Seq.", 1:24)
all.perms
```

B. Let A be the event that only one candidate is interviewed. List the outcomes in "A".

```{r s2.1 q5.b}
A <- all.perms %>% filter(first %in% c(1,2))
A
#############
# Interviews stop after candidate 1 or 2, and only one interview is conducted, so although other sequences exist there are only two outcomes; either candidate 1 or candidate 2 are interviewed first.
#############
```

C. Let B be the event that three candidates are interviewed. List the outcomes in B.

```{r s2.1 q5.c}
B <- all.perms %>% filter(!first %in% c(1,2)) %>% filter(!second %in% c(1,2))
B

# for three candidates to be interviewed, candidates 1 & 2 must be interviewed third or fourth
```
D. Let C be the event that candidate 3 is interviewed.
List the outcomes in C.
```{r s2.1 q5.d}
C <- all.perms %>% filter(first %in% c(3,4)) %>% 
  filter(!(first == 4 & second != 3))
C

# Outcomes where Candidate 3 are interviewed do not have Candidates 1 or 2 preceed them.
```